\documentclass[runningheads]{llncs}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx} % times font instead of cm-modern
\usepackage[margin=1.85cm, paperwidth=6.1in, paperheight=9in, bottom=2.5cm]{geometry}

\usepackage{textgreek}
\usepackage{ucs}
\usepackage{agda}
\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{newunicodechar}
\newunicodechar{⊥}{\ensuremath{\mathnormal\bot}}
\newunicodechar{∷}{\ensuremath{\mathbin{::}}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}

\newunicodechar{≟}{\ensuremath{\stackrel{?}{=}}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≢}{\ensuremath{\not\equiv}}
\newunicodechar{¬}{\ensuremath{\neg}}

\newunicodechar{←}{\ensuremath{\leftarrow}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}

\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{∃}{\ensuremath{\exists}}
%\DeclareUnicodeCharacter{948}{\ensuremath{\delta}}
%\DeclareUnicodeCharacter{955}{\ensuremath{\lambda}}

\usepackage{xcolor}
\usepackage{graphicx}
\pagecolor{white}

\definecolor{body}{RGB}{30, 30, 30}
\color{body}

\begin{document}

\title{Notes on formalizing the STLC}
\author{D. Welch}
\institute{}
\maketitle

\begin{abstract}
Just some notes trying out ``\LaTeX \ literate Agda mode'' as I work through 
this next chapter on the lambda calculus (eventually denotational semantics for
the untyped version of the calculus). Some of the list chapter exercises are
unfinished (including monoids).
\end{abstract}

\section{Intro to the Lambda Calculus} 

The lambda calculus, first published by Alonzo Church in 1932 is a tiny language 
with only three syntactic constructs: variables, abstraction, and (function) 
application. 

The \textit{simply typed lambda calculus} (STLC) is a variant of the lambda 
calculus published in 1940 that adds static typing to the o.g. 1932 untyped 
lambda calculus. It has the three types of constructions previously mentioned 
plus additional syntax to support types and type annotations. We use a 
`Programmable Computable Function' (PCF) style syntax and add operations for 
naturals and recursive function definitions.

Specifically we formalize the base constructs that make up the 
simply-typed lambda calculus: its syntax, small-step semantics, typing, and 
context formation rules. After this a number of properties the language such 
as progress and preservation are stated and proven. The notes may extend the 
language with additional features such as records.

Note: these notes do not present a recommended approach to formalization as 
they eschew a locally nameless representation of terms (via e.g. DeBruijn 
indices). A later section of these notes might look into this. 

\subsection{Imports}

First, we'll need some imports:
\begin{code}%
\>[0]\AgdaKeyword{module}\AgdaSpace{}%
\AgdaModule{stlc}\AgdaSpace{}%
\AgdaKeyword{where}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.Bool}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaDatatype{Bool}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{true}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{false}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaFunction{T}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaFunction{not}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.Empty}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaDatatype{⊥}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaFunction{⊥-elim}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.List}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaDatatype{List}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaOperator{\AgdaInductiveConstructor{\AgdaUnderscore{}∷\AgdaUnderscore{}}}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{[]}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.Nat}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaDatatype{ℕ}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{zero}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{suc}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.Product}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaFunction{∃-syntax}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaOperator{\AgdaFunction{\AgdaUnderscore{}×\AgdaUnderscore{}}}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.String}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaPostulate{String}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaOperator{\AgdaFunction{\AgdaUnderscore{}≟\AgdaUnderscore{}}}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Data.Unit}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaInductiveConstructor{tt}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Relation.Nullary}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaRecord{Dec}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{yes}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{no}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaOperator{\AgdaFunction{¬\AgdaUnderscore{}}}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Relation.Nullary.Decidable}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaFunction{False}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaFunction{toWitnessFalse}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Relation.Nullary.Negation}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaFunction{¬?}\AgdaSymbol{)}\<%
\\
\>[0]\AgdaKeyword{open}\AgdaSpace{}%
\AgdaKeyword{import}\AgdaSpace{}%
\AgdaModule{Relation.Binary.PropositionalEquality}\AgdaSpace{}%
\AgdaKeyword{using}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaOperator{\AgdaDatatype{\AgdaUnderscore{}≡\AgdaUnderscore{}}}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaOperator{\AgdaFunction{\AgdaUnderscore{}≢\AgdaUnderscore{}}}\AgdaSymbol{;}\AgdaSpace{}%
\AgdaInductiveConstructor{refl}\AgdaSymbol{)}\<%
\end{code}

\section{Term Syntax}

Terms have seven constructs. Three are for the core lambda calculus:
\begin{itemize}
\item variables $`x$ 
\item abstractions $λ x \ : \ T$ 
\item applications $L \cdot M$
\end{itemize}

Three are for the naturals, $\mathbb{N}$:
\begin{itemize}
\item zero $`\mathit{zero}$
\item successor $`\mathit{suc} \ M$ 
\item case expressions $\mathit{case} \ L \ [ \mathit{zero} \Rightarrow M | ]$ 
\end{itemize}

There is also one for recursion: 
\begin{itemize}
\item fixpoint $\mu \ x \Rightarrow M$
\end{itemize}

Abstraction is also called lambda abstraction, and is the construct from which 
the calculus takes its name.

With the exception of variables and fixpoints, each term form either constructs
a value of a given type (abstractions yield functions, zero and successor yield 
natural numbers) or deconstructs it (applications use functions, case 
expressions use naturals). This will arise again when we fix rules for assigning
types to terms -- where ctors correspond to intro. rules and deconstructors, 
eliminators.

Here is the syntax of terms in Backus-Naur Form (BNF):
\begin{align*}
L, M, N ::= & `x \ | \ `\lambda x \Rightarrow N \ | \ L \cdot M \ | \\ 
            & `\mathit{zero} \ | \ `\mathit{suc} \ M \ | \ 
            \textit{case} \ L [ \mathit{zero} \Rightarrow M | \mathit{suc} \ x 
            \Rightarrow N ] \ | \ 
            \mu \ x \Rightarrow M 
\end{align*}



\end{document}
